---
title: "stat387"
author: "Mina Mehdinia, Tim Luo, Will McIntosh"
date: "2023-03-01"
output: html_document
---
# Setup

```{r setup, include = F}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)  # Suite of packages incl: dplyr, ggplot2, tidyr, etc.
library(MASS)       # LDA, QDA, OLS, Ridge Regression, Box-Cox, stepAIC, etc...
library(ROCR)       # Precision/Recall/Sens./Spec./ performance plot
library(class)      # KNN, SOM, LVQ
library(e1071)      # Naive Bayesian Classifier, SVM, GKNN, ICA, LCA
library(boot)       # LOOCV, Bootstrap
library(caret)      # Classification/Regression Training
library(randomForest)
library(naivebayes)
# rm(list=ls())
```

```{r read}
german <- read.csv("germancredit.csv", header=T) |> 
  rename(checking = checkingstatus1)

```


```{r remove_redundancy}
set.seed(1)
newgerman
```


# One-Hot Encoding

Below I am making dummy variables (or one-hot encodings) of the categorical variables (WMM - 3/2/23).

```{r}
dummy <- dummyVars(" ~ .", data=german)
newgerman <- data.frame(predict(dummy, newdata = german))
head(newgerman)
```

# Value Counts For Each Class Type

```{r}
Count.Plot <- function(data, column.name) {  
  german.0.default <- data[data$Default == 0, ]
  german.1.default <- data[data$Default == 1, ]
  
  checkingstatus.counts.0 <- table(german.0.default[column.name])
  checkingstatus.counts.1 <- table(german.1.default[column.name])
  
  counts.df.0 <- as.data.frame(checkingstatus.counts.0)
  counts.df.1 <- as.data.frame(checkingstatus.counts.1)
  
  colnames(counts.df.0) <- c(column.name, "count")
  colnames(counts.df.1) <- c(column.name, "count")
  
  counts.df.0 <- counts.df.0 %>% mutate(class = 0)
  counts.df.1 <- counts.df.1 %>% mutate(class = 1)
  
  # Combine the two data frames
  combined_df <- rbind(counts.df.0, counts.df.1)
  
  # Create the plot
  ggplot(combined_df, aes(x = combined_df[,column.name], y = count, fill = as.factor(class))) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(x = column.name, y = "Count", fill = "Class") +
    scale_fill_manual(values = c("#F8766D", "#00BFC4")) +
    theme_minimal()
}
```

## Plotting Checking Status Count Per Class

```{r}
Count.Plot(german, "checkingstatus1")
```

## Plotting History Count Per Class

```{r}
Count.Plot(german, "history")
```

## Plotting Purpose Count Per Class

```{r}
Count.Plot(german, "purpose")
```

## Plotting Savings Count Per Class

```{r}
Count.Plot(german, "savings")
```

## Plotting Employ Count Per Class

```{r}
Count.Plot(german, "employ")
```

# Displaying Duration For Each Class

```{r}
Density.Plot <- function(data, column.name) {
  
  # Create subsets of the dataframe based on the binary class
  df_default0 <- data[data[["Default"]] == 0,]
  df_default1 <- data[data[["Default"]] == 1,]
  
  # Plot the two density plots on the same plot
  ggplot() +
    geom_density(data = df_default0, aes(x = df_default0[,column.name], fill = "Default 0"), alpha = 0.5) +
    geom_density(data = df_default1, aes(x = df_default1[,column.name], fill = "Default 1"), alpha = 0.5) +
    labs(title = paste("Distribution of", column.name, "by Default"),
         x = column.name,
         y = "Density") +
    scale_fill_manual(values = c("#F8766D", "#00BFC4"), name = "Default") +
    theme_minimal()
}
```

# Displaying Amount For Each Class

```{r}
Density.Plot(german, "duration")
```

```{r}
Density.Plot(german, "amount")
```

```{r}
Density.Plot(german, "installment")
```

```{r}
Density.Plot(german, "residence")
```

```{r}
Density.Plot(german, "age")
```

```{r}
Density.Plot(german, "cards")
```

```{r}
Density.Plot(german, "liable")
```


# One-Hot Encoding

Below I am making dummy variables (or one-hot encodings) of the categorical variables.

```{r}
dummy <- dummyVars(" ~ .", data=german)
dummy.german <- data.frame(predict(dummy, newdata = german)) 
head(dummy.german)
```

# RFE

```{r}
RFE <- function(data, num.features=4) {
  # Define the predictor and response variables
  train.X <- data[, !(names(data) %in% c("Default"))]
  train.Y <- as.factor(data[, "Default"])
  
  # Define the control parameters for feature selection
  ctrl <- rfeControl(functions = rfFuncs,
                     method = "cv",
                     number = 10)
  
  # Perform recursive feature elimination using the random forest algorithm
  rf_rfe <- rfe(train.X, train.Y, sizes = c(1:num.features), rfeControl = ctrl)
  
  # Print the results
  print(rf_rfe)
  
  # Plot the results
  plot(rf_rfe, type = c("g", "o"))
}
```

## Running RFE on Dummy-Variabled German Data

```{r}
RFE(german)
```

## Running RFE on Dummy-Variabled German Data

```{r}
RFE(dummy.german)
```

## Running RFE on Dummy-Variabled German Data with Only 2 Variables

```{r}
RFE(dummy.german, num.features=2)
```

## Running RFE on Dummy-Variabled German Data with 10 Variables

```{r}
RFE(dummy.german, num.features=10)
```

## Running RFE on Dummy-Variabled German Data with 20 Variables

```{r}
RFE(dummy.german, num.features=20)
```

# Running PCA Dimensionality Reduction

The first 26 features are statistically significant since they're ab

```{r}
# Standardize the data
dummy.german_std <- scale(dummy.german)

# Perform PCA
german.pca <- prcomp(dummy.german_std, center = TRUE, scale. = TRUE)

# Interpret the results
summary(german.pca)
```

```{r}
# Extract the standard deviations of each principal component
sd <- summary(german.pca)$sdev

# Plot the standard deviations as a line plot
plot(sd, type = "b", xlab = "Principal Component", ylab = "Standard Deviation")
```

# Passing PCA Results in NB

```{r}
# Extract the principal component scores
pc_scores <- predict(german.pca, dummy.german_std)

# Split the data into training and testing sets
set.seed(123) # for reproducibility
train_index <- sample(nrow(dummy.german_std), nrow(dummy.german_std) * 0.7) # 70% for training
train_data <- pc_scores[train_index, ]
train_label <- as.factor(dummy.german_std[train_index,"Default"])
test_data <- pc_scores[-train_index, ]
test_label <- as.factor(dummy.german_std[-train_index,"Default"])

# Train the Naïve Bayes classifier using the training data
model <- naive_bayes(train_data, train_label)

# Predict the test data using the trained model
model.preds <- predict(model, test_data)

# Evaluate the performance of the model
confusionMatrix(model.preds, test_label)
```

# Passing PCA Results in KNN

```{r}
# Extract the principal component scores
pc_scores <- predict(german.pca, dummy.german_std)

# Split the data into training and testing sets
set.seed(123) # for reproducibility
train_index <- sample(nrow(dummy.german_std), nrow(dummy.german_std) * 0.7) # 70% for training
train_data <- pc_scores[train_index, ]
train_label <- as.factor(dummy.german_std[train_index,"Default"])
test_data <- pc_scores[-train_index, ]
test_label <- as.factor(dummy.german_std[-train_index,"Default"])

# Train the KNN classifier using the training data
knn_model <- train(
  x = train_data,
  y = train_label,
  method = "knn",
  trControl = trainControl(method = "cv", number = 10)
)

# Predict the test data using the trained model
knn_pred <- predict(knn_model, newdata = test_data)

# Evaluate the performance of the model
confusionMatrix(model.preds, test_label)
```

