---
title: "stat387"
author: "Mina Mehdinia, Tim Luo, Will McIntosh"
date: "2023-03-01"
output: html_document
---
```{r setup, include = F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)        # LDA, QDA, OLS, Ridge Regression, Box-Cox, stepAIC, etc...
library(ROCR)        # Precision/Recall/Sens./Spec./ performance plot
library(class)       # KNN, SOM, LVQ
library(e1071)       # Naive Bayesian Classifier, SVM, GKNN, ICA, LCA
library(boot)        # LOOCV, Bootstrap
library(caret)       # Classification/Regression Training
library(dplyr)       # For piping and mutation methods
library(ggplot2)     # For plots
library(randomForest)# for RFE
library(naivebayes)  # for NB
library(ROSE)        # for SMOTE
#library(mlbench)
library(fastDummies) # For Get Dummies
library(pROC)        # for PCA
source("Stat387-Proj-HelperFunctions.R")
#library()
# rm(list=ls())
```

# Import Dataset

```{r read}
german <- read.csv("germancredit.csv", header=T)
head(german)[1:7]
```

# Examine Dimensions

```{r}
# Counting Just the rows that did not Default (Good!)
default.0 <- dim(german[german$Default == 0, ])[1]
default.1 <- dim(german[german$Default == 1, ])[1]
paste("The number of those that did not default are", default.0, "and those that did default are", default.1)
```

# Visualizing Dataset Class Imbalance

```{r}
Balance.Plot(german)
```


```{r}
# SMOTE Requires dummy variables
dummy.german <- GetDummies(german)

# Apply SMOTE to balance the dataset
balanced.dummy.german <- ROSE(Default ~ ., data = dummy.german, seed = 123)$data
```

```{r}
Balance.Plot(balanced.dummy.german)
```

# Value Counts For Each Class Type

```{r}
Count.Plot(german, "checkingstatus1")
```

```{r}
Count.Plot(german, "history")
```

```{r}
Count.Plot(german, "purpose")
```

```{r}
Count.Plot(german, "savings")
```

```{r}
Count.Plot(german, "employ")
```

# Plotting Numerical Variable Distributions

```{r}
Density.Plot(german, "duration")
```

```{r}
Density.Plot(german, "amount")
```

```{r}
Density.Plot(german, "installment")
```

```{r}
Density.Plot(german, "residence")
```

```{r}
Density.Plot(german, "age")
```

```{r}
Density.Plot(german, "cards")
```

```{r}
Density.Plot(german, "liable")
```

```{r}
Density.Plot(balanced.dummy.german, "liable")
```

# RFE

```{r}
RFE.4.German <- RFE(german)
```

```{r}
RFE.10.German <- RFE(german, num.features=10)
```

```{r}
RFE.4.Dummy.German <- RFE(dummy.german)
```

```{r}
RFE.2.Dummy.German <- RFE(dummy.german, num.features=2)
```

```{r}
RFE.10.Dummy.German <- RFE(dummy.german, num.features=10)
```

```{r}
RFE.10.Balanced.Dummy.German <- RFE(balanced.dummy.german, num.features = 10)
```

# Running PCA Dimensionality Reduction

```{r}
# According to James et al. (2017), it is generally recommended to standardize
# the dataset before performing PCA, to ensure that the principal components
# accurately capture the variation in the data and the relative importance of
# each variable (p. 373, Chapter 10).

# Perform PCA (PCA Requires Dummy Variables)
dummy.german.std <- StandardizeData(dummy.german)
german.pca <- prcomp(dummy.german.std, center = TRUE, scale. = TRUE)
```

```{r}
# Plot the standard deviations as a line plot
sd <- summary(german.pca)$sdev
plot(sd, type = "b", xlab = "Principle Component", ylab = "Standard Deviation")
```

```{r}
# Plot the Proportion of Variance as a line plot
pov <- german.pca$sdev^2/sum(german.pca$sdev^2)
plot(pov, type = "b", xlab = "Principal Component", ylab = "Proportion of Variance")
```

## Running PCA On the Balanced Dataset

```{r}
# Standardize the data
balanced.dummy.german.std <- StandardizeData(balanced.dummy.german)

# Perform PCA
balanced.dummy.german.std.pca <- prcomp(balanced.dummy.german.std, center=TRUE, scale.=TRUE)
```

```{r}
# Plot the standard deviations as a line plot
sd <- summary(german.balanced.pca)$sdev
plot(sd, type = "b", xlab = "Principle Component", ylab = "Standard Deviation")
```

```{r}
# Plot the "Proportion of Variance" as a line plot
pov <- german.balanced.pca$sdev^2/sum(german.balanced.pca$sdev^2)
plot(pov, type = "b", xlab = "Principal Component", ylab = "Proportion of Variance")
```

# Passing PCA Results in NB

# NB w/ PCA On All Components

```{r}
NB.Model.PCA(german.pca, dummy.german.std)
```

```{r}
NB.Model(dummy.german.std)
```

# NB w/ PCA On First Four Components

```{r}
NB.Model(dummy.german.std)
```

```{r}
german.pca.first.four <- Get.First.n.PCs(german.pca, 4)

```

# NB w/ PCA On First Nine Components

```{r}
german.pca.first.nine <- Get.First.n.PCs(german.pca, 9)
NB.Model.PCA(german.pca.first.nine, dummy.german_std)
```

# NB w/ PCA On First Twenty Components

```{r}
german.pca.first.twenty <- Get.First.n.PCs(german.pca, 20)
NB.Model.PCA(german.pca.first.twenty, dummy.german_std)
```

```{r}
NB.Model(RFE.10.Balanced.Dummy.German)
```

# Passing PCA Results in KNN

```{r}
KNN.Model(dummy.german.std)
```

```{r}
KNN.Model(RFE.4.German)
```

```{r}
KNN.Model(RFE.10.German)
```

```{r}
knn.RFE.10.balanced.dummy.german.fp <- KNN.Model(RFE.10.Balanced.Dummy.German)
nb.RFE.10.balanced.dummy.german.fp <- NB.Model(RFE.10.Balanced.Dummy.German)
```


```{r}
LDA.Model(RFE.4.German)
```

```{r}
QDA.Model(RFE.4.German)
```


```{r}
# Set seed for reproducibility
set.seed(123)

p = 0.7

# Create test/train split
split <- createDataPartition(german$Default, p=p, list=FALSE)

# Subset data into training and testing sets
train.data <- german[split, ]
test.data <- german[-split, ]

print(dim(train.data))
print(dim(test.data))
```

```{r}
RFE.10.dummy.german <- RFE(dummy.german, num.features = 10)
```

```{r}
RFE.10.dummy.german.colnames <- colnames(RFE.10.dummy.german)
RFE.10.dummy.german.colnames
```

```{r}
library(corrplot)

RFE.10.dummy.german.colnames.subset <- dummy.german[RFE.10.dummy.german.colnames]

target <- cor(dummy.german)[,1, drop=FALSE]

corrplot(target, cl.pos="n", method="number")
```

```{r}
df <- data.frame(german)

dictionary <- colname.dict

# Replace values in dataframe using dictionary without changing data types
df[] <- lapply(df, function(x) {
  if(is.character(x)) {
    replace(x, x %in% names(dictionary), dictionary[names(dictionary) %in% x])
  } else {
    x
  }
})
```


```{r}
#df
```

```{r}
#summary(lm(Default~.,df))
```






























